---
title: "03 Baseline Model"
output: html_document
---

```{r}
library(tidyverse)
library(lubridate)
library(forecast)
library(readr)
library(readxl)
```
## Preparing Data
```{r}
# Importing data
daily_data <- read_csv("../data_clean/daily_data.csv")
template   <- read_csv("../data_raw/submission_template.csv")

daily_data <- daily_data %>%
  mutate(date = as.Date(date)) %>%
  arrange(date)

# Transforming data into time series object
ts_load <- ts(
  daily_data$load_daily,
  start = c(year(min(daily_data$date)), yday(min(daily_data$date))),
  frequency = 7
)
```

## Initial Plot
```{r}
TS_Plot <- ggplot(daily_data, aes(x = date, y = load_daily)) +
  geom_line() +
  labs(
    title = "Daily Electricity Load",
    x = "Date",
    y = "Load"
  )
plot(TS_Plot)

# ACF and PACF plots
par(mfrow = c(1, 2))
Acf(daily_data$load_daily, lag.max = 40, main = "ACF of Daily Load")
Pacf(daily_data$load_daily, lag.max = 40, main = "PACF of Daily Load")
par(mfrow = c(1, 1))
```
## Decomposing the time series
```{r}
decompose_load <- decompose(ts_load, type = "additive")
plot(decompose_load)

#Creating non-seasonal load time series because some models can't handle seasonality
deseasonal_load <- seasadj(decompose_load)
```

## Forecasting with Naive methods
```{r}
h <- nrow(template)

# Model 1: Arithmetic mean
MEAN_fit <- meanf(ts_load, h = h)
plot(MEAN_fit)

# Model 2: Naive
NAIVE_fit <- naive(ts_load, h = h)
plot(NAIVE_fit)

# Model 3: Seasonal Naive
SNAIVE_fit <- snaive(ts_load, h = h)
plot(SNAIVE_fit)

#Model summaries
summary(MEAN_fit)
summary(NAIVE_fit)
summary(SNAIVE_fit)
```

## Baseline Model 
We use the seasonal naive model as our baseline because electricity demand exhibits a strong weekly pattern. By setting the frequency to 7, the model captures this seasonality by using the value from the same day in the previous week as the forecast. This provides a simple but meaningful benchmark for more advanced models.
```{r}
baseline_forecast <- as.numeric(SNAIVE_fit$mean)

baseline_model <- template %>%
  mutate(load_daily = baseline_forecast)

head(baseline_model)
```

